---
title: "NYPD_Iraola"
author: "P. Iraola"
date: "2024-04-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NYPD Historical Shooting Report

## Data gathering
These data were pulled from the data.gov catalog. I searched for 'NYPD Shooting Incident Data' and copied the link to the CSV file, which I use in my code below.

## Reading in the data.
First we read in the data, along with the required packages, and then display the data so that we know what it contains.
```{r import, echo=TRUE}
library(tidyverse)
library(stats)
library(nnet)
url = "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
shooting_data <- read_csv(url)
shooting_data
```

## Cleaning the data 
I am interested in comparing perp's age with victim's age, so I clean the data accordingly. I make sure to exclude any missing values, "UKNOWN" values, or "(null)" values.

```{r cleaning, echo=TRUE}
shooting_data <- shooting_data %>%
  select(PERP_AGE_GROUP, VIC_AGE_GROUP) %>%
  filter(!is.na(PERP_AGE_GROUP) & PERP_AGE_GROUP != "UNKNOWN" & PERP_AGE_GROUP != "(null)",
         !is.na(VIC_AGE_GROUP) & VIC_AGE_GROUP != "UNKNOWN" & VIC_AGE_GROUP != "(null)")

shooting_data
```

## Cleaning Part Two
There was still some issues with the data, such as incorrect ages of both victims and perps (most likely typos by whoever was entering the data), so I cleaned it further by filtering for only the following age groups, <18, 18-24, 25-44, 45-64, and 65+.

```{r cleaning2, echo=TRUE}
desired_age_categories <- c("<18", "18-24", "25-44", "45-64", "65+")

shooting_data <- shooting_data %>%
  filter(PERP_AGE_GROUP %in% desired_age_categories &
           VIC_AGE_GROUP %in% desired_age_categories)

shooting_data
```

## Visualizing
Now I want to visualize my data. I decided to use a bar plot to do so.
```{r visualization1, echo=TRUE}
ggplot(shooting_data, aes(x = PERP_AGE_GROUP, fill = VIC_AGE_GROUP)) +
  geom_bar(position = "dodge") +
  labs(x = "Perpetrator Age Group", y = "Count", fill = "Victim Age Group") 
```

## Visualizing Part Two
After looking at my bar plot, I decided I was interested in the two age groups 18-24 and 25-44 as they seem to be where most shootings take place. I created another bar plot that only contains those two categories.
```{r visualization2, echo=TRUE}
filtered_data <- shooting_data %>%
  filter(PERP_AGE_GROUP %in% c("18-24", "25-44"))

# Plot the filtered data
ggplot(filtered_data, aes(x = PERP_AGE_GROUP, fill = VIC_AGE_GROUP)) +
  geom_bar(position = "dodge") +
  labs(x = "Perpetrator Age Group", y = "Count", fill = "Victim Age Group")
```

## Chi Sq Test
I want to see if there is a significant association between the age group of perps and the age group of victims, so I ran a chi sq test of independence.
```{r chisq, echo=TRUE}
chi_sq_test <- chisq.test(table(shooting_data$PERP_AGE_GROUP, shooting_data$VIC_AGE_GROUP))

chi_sq_test
```

## Logistic Regression Model
Now I want to run a logistic regression model on the perp age groups 18-24 and 25-44 to see how well we can predict the victim's age based on the perp's age.
```{r logmodel, echo=TRUE}
# Fit logistic regression model for perpetrators aged 18-24
logit_model_18_24 <- multinom(VIC_AGE_GROUP ~ 1, data = subset(shooting_data, PERP_AGE_GROUP == "18-24"))

# Fit logistic regression model for perpetrators aged 25-44
logit_model_25_44 <- multinom(VIC_AGE_GROUP ~ 1, data = subset(shooting_data, PERP_AGE_GROUP == "25-44"))

logit_model_18_24
logit_model_25_44
```

## Conclusion & Bias
In this analysis I examined the relationship between the age groups of perpetrators and victims involved in shootings in New York. I cleaned the data to focus on certain age groups, visualized the data using bar plots, conducted a chi-square test of independence, and employed logistic regression models to predict victim age based on perp age.

One possible sources of bias is selection bias; this analysis relies on data provided from the NYPD, which may not include all shooting incidents. Unreported shootings could lead to incomplete data and inaccurate conclusions.